---
title: "AssignmentWk2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading and preprocessing the data
###Load the CSV file and convert the date to the appropritate Date format
```{r load}
setwd("C:\\Users\\Said\\Google Drive\\PC\\Education\\Coursera\\Data Science\\Study Materials\\05 Reproducible Research\\Week2\\Assignment")
Data <- read.csv("activity.csv")
Data$date <- as.Date(Data$date, format = "%Y-%m-%d")
```

##What is mean total number of steps taken per day?
### 1. Calculate the total number of steps taken per day
```{r total}
Total_Step <- tapply(Data$steps, Data$date, FUN=sum, na.rm=TRUE)
```

### 2. Make a histogram of the total number of steps taken each day
```{r hist}
hist(Total_Step)
```

### 3. Mean and median number of steps taken each day
```{r mean}
Mean_Step <- mean(Total_Step, na.rm = TRUE)
Mean_Step
Median_Step <- median(Total_Step, na.rm = TRUE)
Median_Step
```

##What is the average daily activity pattern?
### 1. Time series plot of the average number of steps taken
```{r TimeSeriesPlot}
Step_Interval <- aggregate(x=list(steps = Data$steps), by  = list(interval = Data$interval), FUN=mean, na.rm=TRUE)
plot (Step_Interval, type="l")
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r MaxStep}
Step_Interval[which.max(Step_Interval$steps),]
```

## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r Missing}
NA_Data <- is.na(Data$steps)
sum(NA_Data)
```

### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc
```{r Fill_Missing}
Step_Interval_DF <- as.data.frame(Step_Interval)
NewData <- merge (Data,Step_Interval_DF, by = "interval")
```

### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r Impute}
NewData$steps.x[is.na(NewData$steps.x)] = NewData$steps.y[is.na(NewData$steps.x)]
```

### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r Hist_Impute}
Total_Impute_Step <- tapply(NewData$steps.x, NewData$date, FUN=sum, na.rm=TRUE)
hist(Total_Impute_Step)
Mean_Impute_Step <- mean(Total_Impute_Step, na.rm = TRUE)
Mean_Impute_Step
Median_Impute_Step <- median(Total_Impute_Step, na.rm = TRUE)
Median_Impute_Step
```
The values of Mean and Median increase after imputing the missing value. Mean and Median value increases when the missing values are filled up with more large numbers.

## Are there differences in activity patterns between weekdays and weekends?

## 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r Weekday_Factor}
NewData$day <- weekdays(NewData$date)
NewData$day[NewData$day=="Monday"|NewData$day=="Tuesday"|NewData$day=="Wednesday"|NewData$day=="Thursday"|NewData$day=="Friday"] <- "Weekday"
NewData$day[NewData$day=="Saturday"|NewData$day=="Sunday"] <- "Weekend"
```

## 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)  
```{r TimeSeriesPlot_Weekday}
library(ggplot2)
Step_Interval_Weekday <- aggregate(steps.x ~ interval + day, data = NewData, mean)
ggplot(Step_Interval_Weekday, aes(interval, steps.x)) + geom_line() + facet_grid(day ~ .) + 
   ylab("Steps")

```
